---
title: "Life Expectancy Example"
author: "Jeffrey B. Arnold"
date: "05/12/2015"
output: html_document
---

This example will use data from the Barro and Lee dataset to analyze 
life expectancy. 
This is a pedagogical example, and does not represent a sophisticated epidemiological model.
Data are 138 countries in 1985.

```{r barro_csv}
barro_raw <- read.csv("http://pols503.github.io/pols_503_sp15/data/barro.csv")
```

Subset of barro data with only the variables we need, and dropping missing observations.
```{r barro}
barro <- barro_raw %>%
  select(lifeexp, school, gdpcap85,
         civlib5, wartime) %>%
  na.omit

glimpse(barro)
```

gdpcap85

:   Per capita GDP in 1985, thousands of international dollars

school

:   Average years of education

civlib5

:   low = 1 to high = 7 scale of civil liberties

wartime

:   Percent of recent history spent at war


Regression of Life Expectancy on `gdpcap85`, `school`, `civlib5`, and `wartime`.

```{r mod1}
mod1 <- lm(lifeexp ~ gdpcap85 + school + civlib5 + wartime,
           data = barro)
```

```{r mod1}
mod2 <- lm(lifeexp ~ log(gdpcap85) + log(school) + civlib5 + wartime,
           data = barro)
```

## Comparing the two models

```{r mod1_glance}
glance(mod1)
```

```{r}
glance(mod2)
```

```{r}
library(boot)

mod1_glm <- glm(lifeexp ~ gdpcap85 + school + civlib5 + wartime, data = barro, na.action="na.omit")

# Note the use of glm rather than lm
cv_err_K10 <- cv.glm(barro, mod1_glm, K=10)

# This runs the cross-validation
cv_err_K10$delta

```
And the logged model
```{r}
mod2_glm <- glm(lifeexp ~ log(gdpcap85) + log(school) + civlib5 + wartime, data = barro, na.action = "na.omit")

cv_err_K10 <- cv.glm(barro, mod2_glm, K=10)

# This runs the cross-validation
cv_err_K10$delta
```

Leave-one-out cross-validation
```{r}
cv.glm(barro, mod1_glm)$delta
cv.glm(barro, mod2_glm)$delta

```

```{r}
augment(lm(lifeexp ~ log(gdpcap85) + log(school) + civlib5 + wartime,
        data = barro))
```


## Predictive Comparisons by Simulation and Bootstrap

### Simulating Coefficients from Asymptotic Normal

We want to compare data 

Use the method of King, Tomz, et al.

Draw simulation
$$
\tilde\beta \sim N(\hat\beta, V(\hat\beta))
$$
This relies on the CLT result that $\hat\beta$ approx multivariate normal as $n \to \infty$.
```{r}
n <- 1024
simbetas <- MASS::mvrnorm(n, coef(mod1), vcov(mod2))
```

For `log(gdp)` compare it at its mean versus 1 standard deviation
```{r}
barro_low <- summarize(na.omit(barro),
                      gdpcap85 = exp(mean(log(gdpcap85))),
                      school = exp(mean(log(school))),
                      civlib5 = mean(civlib5),
                      wartime = mean(wartime))
barro_high <- summarize(na.omit(barro),
                       gdpcap85 = exp(mean(log(gdpcap85)) + sd(log(gdpcap85))),
                       school = exp(mean(log(school))),
                       civlib5 = mean(civlib5),
                       wartime = mean(wartime))
```



Simulate and calculate a standard error around the difference
```{r}
xlow <- model.matrix(~ log(gdpcap85) + log(school)
                      + civlib5 + wartime, data = barro_low)
xhigh <- model.matrix(~ log(gdpcap85) + log(school)
                      + civlib5 + wartime, data = barro_high)
diffs <- rep(NA, nrow(simbetas))
for (i in 1:nrow(simbetas)) {
  betas <- simbetas[i, ]
  diffs[i] <- xhigh %*% betas - xlow %*% betas
}

```

This could be done much easier with **Zelig** or **simcf**.

## Bootstrapping

The function `Boot` from **car** can be used to bootstrap coefficients.
```{r}
Boot(mod2)
```

The **boot** package has more powerful and general functions.

Bootstrapping can be done manually with `dplyr` functions:
```{r}
simulations <- list()
for (i in 1:1024) {
  # Randomly resample data
  .data <- sample_frac(barro, replace = TRUE)
  # Run a new regression
  mod <- lm(lifeexp ~ log(gdpcap85) + log(school) + civlib5 + wartime, data = .data)
  # Predict
  lower <- predict(mod, newdata = barro_low)
  higher <- predict(mod, newdata = barro_high)
  simulations[[i]] <-
    data.frame(diff = higher - lower)
}
sims <- bind_rows(simulations)

quantile(sims$diff, c(0.025, 0.975))
```

This could also be done using the `bootstrap` function in **broom**.

