---
title: "Panel Data"
author: "Jeffrey B. Arnold"
date: "06/01/2015"
output: html_document
---

This worked example introduces estimation in panel data in R using example on partisan governments and economics measures for a panel of OECD states.

# Setup

This tutor
```{r message = FALSE}
library("lmtest")
library("car")
library("dplyr")
library("ggplot2")
```

The panel data specific packages used are **plm** and **pcse**:
```{r message  = FALSE}
library("plm")
library("pcse")
```



# Data
 
This example data was the data from Garrett (1989) *Partisan Politics in the Global Economy*.
The analyses are similar (but not the same) as those run in King, Wittenberg and Tomz (2000), Bailey and Katz, and Beck and Katz (2011). 
The replication data is from Beck and Katz (2011).


```{r}
garrett1998 <- read.csv("../data/garrett1998.csv", stringsAsFactors = FALSE)
```

The data consist of `r nrow(garrett1998)` observations for `r length(unique(garrett1998)$county)` countries for `r length(unique(garrett1998)$year)` years (`r min(garrett1998$year)`--`r(max(garrett1998$year))`.

unem

:    unemployment

inf

:    inflation

gdp

:    Growth in GDP

leftlab

:    Political power of left parties in the government. Index from 0--5.

corp

:    A measure of corporatism or how encompassing the labor movement is. Index from 0--4.

capmob

:    Capital mobility. Index from 0--4.

oild

:    Oil dependency

demand

:    OECD demand (growth rate in all OECD countries)



Effect of the left parties, corporatism and the interaction of the two.

## Dummy variable models

Pooled time-series, cross-section

```{r}
mod1 <- lm(gdp ~ corp * leftlab + demand + oild + trade, data = garrett1998)
```

Controlling for country

```{r}
mod2 <- lm(gdp ~ corp * leftlab + demand + oild + trade + country,
           data = garrett1998)
```


Controlling for country and year

```{r}
mod3 <- lm(gdp ~ corp * leftlab + demand + oild + trade + country + year,
           data = garrett1998)
```

Lagged dependent variable

```{r}
mod_ldv <- lm(gdp ~ gdpl + corp * leftlab + demand + oild + trade + year,
              data = garrett1998)
mod_ldv
```

Lagged dependent variable plus PCSE
```{r}
pcse(mod_ldv, groupN = garrett1998$country, groupT = garrett1998$year)
```


# Panel Models using plm

The functions in **plm** need to know which variables correspond to id and time in the data frame. 
So use the function `pdata.frame` to create a data frame friendly with the **plm** functions.
```{r}
pgarrett <- pdata.frame(garrett1998, index = c("country", "year"))
```

Within (fixed-effects) model

```{r}

mod_within <- plm(gdp ~ corp * leftlab + demand + oild + trade + year,
                  data = pgarrett, model = "within")
summary(mod_within)
```

Recover the fixed effects 

```{r}
fixef(mod_within)
```

Random effects model

```{r}
mod_re <- plm(gdp ~ corp * leftlab + demand + oild + trade,
                  data = pgarrett, model = "random")
summary(mod_re)
```

Compare the random effects model with the fixed effects model using a Hausman test
```{r}
phtest(mod_within, mod_re)
```
If we can reject the null, then we should use the fixed effects estimator.




