---
title: "POLS 503: Problem Set 3"
author: Anonymous
date: "April 30, 2015"
output:
  html_document: default
---

# Setup

This homework depends on the following packages.
```{r load, message = FALSE}
library("ggplot2")
library("dplyr")
library("broom")
library("assertthat")
library("tidyr")
library("pols503")
```
This homework uses functions from the distributed file `hw3-functions.R`.
```{r}
source("hw3-functions.R")
```

This code chunk sets default code chunk settings that will make your simulations quicker and reproducible: see the [R Markdown Reference Guide](http://www.rstudio.com/resources/cheatsheets/).
In particular, it ensures that knitr will cache results so each time you compile,  you do not need to rerun simuations if things do not change. 
```{r}
library("knitr")
opts_chunk$set(cache = TRUE,
					    autodep = TRUE,
					    cache.extra = rand_seed)
```
If code chunks do not evaluate, just delete that chunk or set `cache = FALSE` to force every.
Finally, we would like to ensure that we get the same results each time it is run even though we are drawing random numbers.
The way to do this is to set a seed for the random number generator.
```{r seed}
set.seed(11346170)
```

# Problems

All simulations will use 1,024 iterations.
```{r}
iter <- 1024
```

## Problem 1: Linear Normal Model with Homoskedasticity

My answer here ...

## Problem 2: Correlated Covariates

My answer here ...

## Problem 3: Collinearity

My answer here ...

## Problem 4: P-Values, Type I and II Errors

My answer here ...

## Problem 5: Omitted Variable Bias

My answer here ...

## Problem 6: Heteroskedasticity

My answer here .. 
