---
title: "POLS 503: Problem Set 4"
author: "Christopher Adolph, Jeffrey B. Arnold"
date: "May 15, 2015"
---

$$
\DeclareMathOperator{\logit}{logit}
$$

- 2015-05-19: Added logit function. Add note on dealing with data transformations in sprinters data.
- 2015-05-20: Added function using **simcf** functions for simulation.


```{r load,message = FALSE, echo = FALSE}
library("MASS")
library("ggplot2")
library("dplyr")
library("broom")
library("tidyr")
```


# Instructions and Introduction

1. Create a new R project for this homework named `hw4` and load that project.
2. Download [hw4-functions.R](hw4-functions.R) and save in to your project directory.
3. Submit a zipped file of the directory with your R project through Canvas.
    This should contain all the materials for another person to run your R Markdown file  This should contain:

    - The R project (`.Rproj`) file.
	  - The R Markdown document (`.Rmd`) of your analyses.
	  - An HTML document (`.html`) compiled from your R Markdown document.
	  - Any data or other files needed to run the analyses in your R Markdown document.

4. Your R Markdown file should follow the guidance from [here](http://pols503.github.io/pols_503_sp15/r_markdown_assignments.html) and your R code should follow the guidelines [here](http://pols503.github.io/pols_503_sp15/r_best_practices.html).
   **You should not include many, if any, comments in your code chunks**. 
   Your discussion should be included directly in the document.
   
5. Turn in a paper copy of the document compiled from the analyses at lab.
6. You can work together on this but you should each turn in your own assignments and write up your work separately.
Include the names of your collaborators on your assignment.

Something which will be of use to you later is the logit function.
The logit function is defined as
$$
\logit(x) = \log \left(\frac{x}{1-x} \right) = \log(x) - \log(1 - x)
$$
where $x \in (0, 1)$. It takes a value between 0 and 1 and returns a value between $-\infty$ and $\infty$.
```{r}
logit <- function(x) log(x) - log(1 - x)
```

# Problems {#problems}

## Showing Confidence

We will revisit the sprinters data we considered in Problem Set 2.
```{r}
sprinters <- read.csv("http://UW-POLS503.github.io/pols_503_sp15/data/sprinters.csv")
```

a. Estimate the model
    $$
    \begin{aligned}[t]
    \mathtt{finish}_i &= \beta_0 + \beta_1 \mathtt{year}_i + \beta_2 \mathtt{women}_i + \beta_3 \mathtt{women}_i \times \mathtt{year}_i  + \epsilon_i
    \end{aligned}
    $$
    Create a plot of the fitted values with confidence with respect to `year`, grouped by `women`.
    Do not use `geom_smooth()`.

b. Create the same plot as in the previous part for the model
    $$
    \begin{aligned}[t]
    \log (\mathtt{finish}_i) &= \beta_0 + \beta_1 \mathtt{year}_i + \beta_2 \mathtt{women}_i + \beta_3 \mathtt{year}_i \times \mathtt{women}_i + \epsilon_i
    \end{aligned}
    $$
    Be sure to explain in words how this specification differs from the one used in part a.

c. Rerun the analysis and recreate the plot, adding confidence intervals, for the
model:
    $$
    \begin{aligned}[t]
    \mathtt{Finish}_i &= \beta_0 + \beta_1 \mathtt{year}_i + \beta_2 \mathtt{women}_i + \beta_3 \mathtt{year}_i × \mathtt{women}_i \\
	& + \beta_4 \mathtt{year}_i^2 + \beta_5 \mathtt{year}_i^2 × \mathtt{women}_i + \epsilon_i
    \end{aligned}
    $$
    Be sure to explain in words how this specification differs from the ones used in part a. and b.
d. Compare the visual fit of these models to the data within the observed period.
    Which do you find plausible fits?
e. Do these models have different predictions for the Olympics of 2156? (Hint:
    extending your plots to go up to 2156 is an easy way to see this.) Why or why
    not?
f. Now create a new variable, the ratio of men’s time to women’s time in each
   year.
   Logit-transform this variable and regress it on year. Plot the results,    with confidence intervals, on the scale of the ratio men’s time to women’s time (i.e., transform it back from logit).
   Does this approach make any assumptions about men’s times or women’s times that might be problematic?
   *Hint:* You will need to manipulate the data frame for this. Your new data frame will have years as the unit of observation rather than (year, sex). You can do this using manually using `filter` and other post-processing. Alternatively, look at the **tidyr** function `spread`.

## Model Selection: Oil & Democracy

For this problem, we will use a cleaned-up version of the dataset of Michael Ross, "Does Oil Hinder Democracy?" *World Politics*, 2001.
That paper estimated a time series cross-section model of Polity scores regressed on oil exports and a battery of controls.
In this problem, we will focus on a single cross-section, and instead focus on model fitting.

Load and pre-process this data using the following code:
```{r}
rossoil <- read.csv("http://UW-POLS503.github.io/pols_503_sp15/data/rossoildata.csv") %>%
   arrange(id1, year) %>%
   group_by(id1) %>%
   mutate(oilL5 = lag(wdr123, 5) / 100,
          metalL5 = lag(wdr313, 5) / 100,
          GDPpcL5 = lag(wdr135, 5) / 100,
          islam = islam / 100) %>%
   filter(year == 1990) %>%
   select(regime1, oilL5, metalL5, GDPpcL5, islam, oecd, cty_name, id, id1) %>%
   na.omit() %>%
   ungroup()
```

The above code:

- `lag()` calculates the lag values of `oil`, `metal`, and `GDPcap`. The second argument of `5`, means that it calculates a lag of 5.[^1]
- `group_by` ensures that the lag values are only calculated within each country.
- Keeps only observations from 1995 and a subset of variables
- Omits missing values with `na.omit()`.
- `ungroup()` ensures that the data is no longer grouped by `id`.
  If you try to use `summarize()` while the data is still grouped by `id`, you would
  not get the results that you thought you would.

[^1]: The original `ross95` data has separate variables with the lagged values, but I wanted to show how to calculate lags with `dplyr`.

A description of the included variables follows:

| Variable | Description |
|:--|:--|
| `regime1` |    1–10 scale increasing in democracy; computed from Polity components |
| `oilL5` |      Fuel exports as a proportion of GDP, lagged 5 years |
| `metalL5` |    Ore and mineral exports as a proportion of GDP, lagged 5 years |
| `GDPcapL5` |    per capita GDP in PPP dollars, lagged 5 years |
| `islam` |      Muslims as a proportion of population, 1970 data |
| `oecd` |       Dummy for rich industrialized countries |
| `cty_name` |   The name of the country observed |
| `id` |         A three character abbreviation of the country name |
| `id1` |        A numeric country code |



a. Estimate a linear regression of `regime1` on `oilL5`, `metalL5`, `GDPcapL5`, `islam`, and `oecd`.
  Record the standard error of the regression, and calculate the expected change in `regime1` given a change in `oilL5` from the 50th percentile to the 95th percentile of the fully observed data, all else equal.
b. Using the residuals from the regression in part a., create the following diagnostic plots:
    
    1. Plot the residuals against the fitted values
    2. Plot the residuals against each covariate
    3. Plot the studentized residuals against the standardized hat values.
    4. Calculate the heteroskedasticity consistent standard errors and compare to the classical standard errors. Use the **car** function `hccm`.
    
    What do these diagnostics tell you about the presence of heteroskedasticity, specification error, and outliers?
    
c. Rerun the regression using either log or logit transformations on any covariates you see fit.
    You will likely run several specifications.
    In each run, record the standard error of the regression, and the expected change in `regime1` given a change in `oilL5` from the 50th percentile to the 95th percentile of the fully observed data.
    See the appendix for some tips and warnings about transforming these data, though.
d. How much substantive difference does finding the best model make?
    Be specific and concrete; i.e., show what each model does.
    I’m asking for a more detailed answer than you usually see in articles.
    How much substantive doubt is there in the result if we are not sure which of the models you fit is the  "right" one?
e. Which model of those you have estimated do you trust most, and why?
    What other problems in the specification or estimation method remain unaddressed by our efforts?
    
# Appendix:  How Do I Log a Covariate with Zeros?

*Christopher Adolph*

If you try to log or logit transform a covariate $x$ with observed
zeros, you will discover a problem: you can't log a zero!
A common (but wrong) "solution" is to add a small amount to the zeros
(e.g., 0.1 or 0.001, etc.).
It turns out that you can introduce substantial large bias in your $\hat \beta$s by choosing different tiny amounts to add to your 0s: logging small numbers spreads those numbers out over a huge range.
Adding 0.001 before logging a variable is not very different from subtracting 10,000 from an unlogged variable!
So don't ever do this, even as a first try.

## A Solution: the logBound and logitBound Transformations

A better solution that avoids arbitrary assumptions and bias is to
"dummy out" the zeros before logging. 
This procedure treats the zero cases as *sui generis*:  they are uniquely different from the rest of our cases,
and we estimate the way in which they are different through a
separate parameter.  We end up with two variables on the right-hand
side: an indicator of whether $x_i=0$, and the log (or logit) of
$x_i$ in those cases where $x_i \ne 0$.  That is, if you want to
regress $y$ on $\textrm{log}(x)$ but $x$ contains $0$s, estimate this
regression:
$$
y_i = \beta_0 + \beta_1 I(x_i>0) + \beta_2 \log'(x_i) + \epsilon_i
$$
where $I(\cdot)$ is an indicator function and $\log'(\cdot)$ is defined as:
$$
\log'(x) =
\begin{cases}
0 &\text{if $\quad x \le 0$} \\
\log(x) & \text{if $x > 0$} 
\end{cases}
$$

If we suppose that $x_i$ is the number of cigarettes person $i$ smokes
per day, and $y_i$ is $i$'s probability of getting lung cancer, the
specification makes sense:  people who currently smoke even a little bit
likely have a discretely higher chance of lung cancer than
non-smokers, while the amount a smoker smokes may increase cancer
probabilities but with diminishing marginal risk.

As you might imagine, the logic of equations 1 and 2
changes slightly if $x$ needs to be logit transformed.  Recall that the logit
transformation,
$$
\logit(x) = \log\left( \frac{x}{1-x} \right),
$$
\noindent fails if $x$ is not between 0 and 1, so we need to dummy out $x_i \ge 1$ and $x_i \le 0$ separately:
$$
y_i = \beta_0 + \beta_1 I(x_i>0) + \beta_2 I(x_i \ge 1) + \beta_3 \logit'(x_i) + \epsilon_i
$$
where $I(\cdot)$ is an indicator function and $\logit'(\cdot)$ is defined as:
$$
\logit'(x) =
\begin{cases}
0 &\mathrm{if} \quad x \le 0 \\
\log \left( \frac{x}{1-x} \right)&
    \mathrm{if} \quad 0<x<1 \\ 
0 &\mathrm{if} \quad x \ge 1
\end{cases}
$$
Note that we will only need all three pieces of Equation 4 if the covariate to be logit transformed contains **both** 0s and 1s; should either extreme be missing, we need only add one dummy variable to the specification.

The `logBound()` and `logitBound()` functions are included in `hw4-functions.R`.
You can load these functions with the following code (provided you downloaded `hw4-functions.R` and put it in your package directory):
```{r}
source("http://UW-POLS503.github.io/pols_503_sp15/hw/hw4-functions.R")
```
After loading these functions, the above regression is as simple as:
```{r eval=FALSE}
res <- lm(y ~ logBound(x), data)
```
You can compare goodness of fit as usual.
Moreover, you can use this technique on the right-hand side of any regression-like model, not just least squares regression.

## Interpretation of results

Take special care in interpreting models in models with `logBound(x)` or `logitBound(x)` in the model formula.
In setting up a hypothetical scenario for post-estimation prediction, make sure both the dummy term and the log term are set consistent with each other.
For example, if the dummy is set to 0, the log must also be zero.
And if the log is set to something other than 0, the dummy must be set to 1.
Otherwise, you are asking the model to predict a logically impossible scenario; e.g., asking what happens when someone both smokes zero cigarettes and smokes twenty cigarettes in the same day.

I recommend either calculating the predicted values of `regime1` "by hand", or using the **simcf** package, as illustrated below.
Our old friend `predict()` is very unlikely to return results for models including these terms, though if it does return an answer it will agree with other methods.

# Appendix: Example Code

This section provides some example code that will make it easier to answer the homework using simulation from **simcf**.

If you haven't already installed **simcf**, you will need to install it. 
You will need to have **devtools** installed.
```{r eval = FALSE}
library("devtools")
install_github("chrisadolph/tile-simcf", subdir = "tile")
install_github("chrisadolph/tile-simcf", subdir = "simcf")  
```

The following function, `predOil` applies simulation to the particular problem in this problem set.
The arguments of `predOil` are:

res

:    The object returned by `lm()`.

formula

:    Formula used in the initial call to `lm()` that created `res`.

data

:    The data used in the initial call to `lm()` that created `res`.

sims

:    The number of simulations to run.

ci

:    A number between 0 and 1. The confidence level to use in confidence interval. 

This will return a `data.frame` object with columns

- `pe`: Point estimate
- `lower`: Lower bound of the confidence interval
- `upper`: Upper bound of the confidence interval

```{r message = FALSE, eval = FALSE}
library("simcf")
predOil <- function(res, formula, data, sims = 1000, ci = 0.95) {
  pe <- res$coefficients
  vc <- vcov(res)
  simbetas <- mvrnorm(sims, pe, vc)
  xscen <- cfMake(formula, data = data, nscen = 1)
  xscen <- cfChange(xscen, "oilL5",
                    x = quantile(data$oilL5, probs = 0.95),
                    xpre = quantile(data$oilL5, probs = 0.5),
                    scen = 1)
  as.data.frame(linearsimfd(xscen, simbetas, ci = ci))
}
```
An example call to this function.
```{r eval = FALSE}
mod0_formula <- regime1 ~ oilL5 + metalL5 + GDPpcL5 + logitBound(islam) + oecd
mod0 <- lm(mod0_formula, data = rossoil)
pred0_oil <- predOil(mod0, mod0_formula, rossoil)

```

It may be useful to combine the estimates from several runs using the function `bind_rows` from **dplyr**.

* * *

Derived from of Christopher Adolph, "Problem Set 4", *POLS/CSSS 503*, University of Washington, Spring 2014. <http://faculty.washington.edu/cadolph/503/503hw4.pdf>. Used with permission.

