---
title: 'Lab 2: Graphing and Data Manipulation'
author: "Jeffrey B. Arnold and Sergio Garcia-Rios"
date: "April 10, 2015"
output: html_document
---

We will be using the following libraries in this `dplyr` and `ggplot2` so make sure to install and load those libraries:

```{r, libraries, warning = FALSE, message = FALSE}
library("dplyr")
library("ggplot2")
library("gapminder")
```


# Meet the mighty `dplyr`

**dplyr** aims to provide a function for each basic verb of data manipulating:

-  `filter()` and `slice()`
-  `arrange()`
-  `select()` and `rename()`
-  `distinct()`
-  `mutate()` and `transmute()`
-  `summarise()`
-  `sample_n()` and `sample_frac()`

## Load the Data

We will be using the gapminder data again.
The 

Let's take a look at the structure of this dataset using `str()` which gives us useful information about the internal structure of any `R` object.
```{r}
str(gdf)
```

We have previously used ``head()`` which is useful to get a quick view of your data but this time we are trying something different.

We are going to use `tbl_df()` which is a wrapper around a data frame that won’t accidentally print a lot of data to the screen

```{r}
gtbl <- tbl_df(gdf)
gtbl
```

A `tbl_df` is basically an improved data.frame, for which dplyr provides nice methods for high-level inspection.

We can also use `glimpse`
```{r}
glimpse(gtbl)
```

## Exploring our Data

We are ready to begin exploring our data set in more depth.

We want to explore the relationship between life expectancy and GDP.
Let's use some `dplyr` verbs to explore our data.
For you Stata users missing "if statements" let's begin with `filter()`

```{r}
filter(gtbl, lifeExp < 29)
filter(gtbl, country == "Rwanda")
filter(gtbl, country %in% c("Rwanda", "Afghanistan"))
```


## %>%

Before we go any further, we should exploit the new pipe operator that `dplyr` imports from the **magrittr** package. This is going to change your data analytical life.

Here’s what it looks like: `%>%`
```{r}
gdf %>% head()
```

This is equivalent to `head(gdf)`. This pipe operator takes the thing on the left-hand-side and pipes it into the function call on the right-hand-side - literally, drops it in as the first argument.

Of course, you can still specify other arguments to this function! To see the first 3 rows of Gapminder, we could say:
```{r}
gdf %>% head(3)
```

# Use filter() to subset data row-wise.
filter() takes logical expressions and returns the rows for which all are TRUE.

```{r}
filter(gtbl, lifeExp>40) %>%
  head(3)
```


## Use `select()` to subset the data on variables or columns.

Most of the times we don't need to see all the variables and are often interested in just a few of them. Here’s a conventional call:
```{r}
select(gtbl, year, lifeExp)
```

# Missing Data

# Challenge

Using a combination of `filter`, `select` and and `head` show only year and life expectancy of Cambodia.for the first two observations


```{r}
gtbl %>%
  filter(country == "Cambodia") %>%
  select(year, lifeExp) %>%
  head(2)
```

# Use `mutate()` to add (generate?) new variables

Imagine we wanted to recover each country’s GDP. We do have data for population and GDP per capita. what do we do?

-  Yes we multiply, let's create a new variable called GDP that brings back the gross amount

```{r}
gtbl <- gtbl %>%
mutate(gdp = pop * gdpPercap)


gtbl %>% glimpse #let's take a look
```

So... GDP is almost useless because it doesn't give a base line and that is why we often use per capita, but a baseline is often more useful, how about comparing it to another country sat, USA? Yes USA, USA, USA!

Let's  create first an object containing only US data, we use `filter` here
```{r}
just_usa <- gtbl %>% filter(country == "United States")

```

Remember our friend `match` from that 501 R homework, Yeah I don't remember either... Well, here it is again!
```{r}
gtbl <- gtbl %>%
  mutate(usa = just_usa$gdpPercap[match(year, just_usa$year)],
         gdpPercapRel = gdpPercap / usa)

gtbl

gtbl %>%
  select(gdpPercapRel) %>%
  summary
```

Nice, now we can do something like this:

Look at the GDP per capita  of Mexico and Canada relative to US by year
```{r}
gtbl %>%filter(country==c("United States", "Canada", "Mexico")) %>%
  select(country, year, gdpPercap, usa, gdpPercapRel)
```

Great! Mexico is way poorer in relation to US.



## `arrange` because the world is not always ordered the way we want it
```{r}
gtbl %>%
  arrange(year, country)


gtbl %>%
  filter(year == 2007) %>%
  arrange(lifeExp)

```


## Challenge


What about life expectancy? Create a relative to life expectancy variable, compare the three NAFTA countries US, Canada and Mexico


```{r}
# Challenge here
```

WOW! Mexicans are relatively poor but not dying, Great!

# Plotting with ggplot2

We will be using the graphics package **ggplot2**, which is one of the most popular, but it is only one of several graphics packages in R.[^1]

[^1]: You may encounter these other packages in other classes, or code samples online.

	- base graphics: included with R. Functions like `plot`, `barplot`, `hist`. See http://www.statmethods.net/graphs/index.html
	- lattice graphics: Functions including `barchart`, `densityplot`, `dotplot`, `xyplot`, `histogram`. See http://www.statmethods.net/advgraphs/trellis.html
	- Other graphics packages designed for the web: [ggvis](http://ggvis.rstudio.com/), [rcharts](http://rcharts.io/), [plotly](https://plot.ly/)
	- [tile](http://faculty.washington.edu/cadolph/?page=60) Chris Adolph's graphics package.

## Grammar of graphics

Unlike many other graphics systems, functions in **ggplot2** do not correspond to separate types of graphs.
There are not scatterplot, histogram, or line chart functions per se.
Instead plots are built up from component functions.

1. Data
2. Aesthetics: Maps variables in the data to visual properties: position, color, size, shape, line type ...
3. Geometric objects: The specific shapes that are drawn: points, lines, 
4. scales: How variables values map to "computer" values.
5. stat: summarize or transform the data. e.g. bin data and count in histogram; run a regression to get a line.
5. facet: create mini-plots of data subsets


Let's continue using the gapminder data, take another look at it
```{r}
glimpse(gtbl)
```

Great, let the plotting begin:

```{r}
 # ggplot(gtbl, aes(x = gdpPercap, y = lifeExp))
```
Nothing to plot yet! This just initializes the gg, it is better if you assign it to an object, `p` is good enough
```{r}
p <- ggplot(gtbl, aes(x = gdpPercap, y = lifeExp))
```

No we can add `geoms`
```{r}
p + geom_point()
```


That look okay but it would probably look be better if we log transform
```{r}
p_l <- ggplot(gtbl, aes(x = log10(gdpPercap), y = lifeExp))
p_l + geom_point()
```

A better way to log transform
```{r}

p + geom_point() + scale_x_log10()

```

Let's make that stick
```{r}
p <- p + scale_x_log10()

```

Common workflow: gradually build up the plot you want,  re-define the object 'p' as you develop "keeper" commands.
Note that in the reassigning we excluded the `geom`.


Now, let continent variable to aesthetic color
```{r}
p + geom_point(aes(color = continent))

```

In full detail, up to now:
```{r}
ggplot(gtbl, aes(x = gdpPercap, y = lifeExp, color = continent)) +
  geom_point() + scale_x_log10()
```

Let's address over-plotting: SET alpha transparency and size to a value
```{r}
p + geom_point(alpha = (1/3), size = 3)

```

Add now a fitted curve or line
```{r}
p + geom_point() + geom_smooth()
p + geom_point() + geom_smooth(lwd = 2, se = FALSE)

p +  geom_smooth(lwd = 1, se = FALSE, method = "lm") + geom_point()

```
That's great but i actually think want revive our interest in continents!
```{r}

p + aes(color = continent) + geom_point() + geom_smooth(lwd = 3, se = FALSE)
```
Facetting: another way to exploit a factor
```{r}
p + geom_point(alpha = (1/3), size = 3) + facet_wrap(~ continent)

# Still want lines? Let's add them
p + geom_point(alpha = (1/3), size = 3) + facet_wrap(~ continent) +
  geom_smooth(lwd = 2, se = FALSE)
```

## Challenge:

* plot lifeExp against year
* make mini-plots, split out by continent
* add a fitted smooth and/or linear regression, w/ or w/o faceting


## Saving data

You can `ggsave`



```{r}
## CHALLENGE HERE


# plot lifeExp against year
# y <- ggplot(gDat, aes(x = year, y = lifeExp)) + geom_point()

# make mini-plots, split out by continent
# y + facet_wrap(~ continent)

# add a fitted smooth and/or linear regression, w/ or w/o facetting
#y + geom_smooth(se = FALSE, lwd = 2) +
#  geom_smooth(se = FALSE, method ="lm", color = "orange", lwd = 2)

# y + geom_smooth(se = FALSE, lwd = 2) +
#  facet_wrap(~ continent)

```

What if I am only interrested in the US?

sadly, ggplot() does not have a 'subset =' argument
so do that 'on the fly' with subset(..., subset = ...)

```{r}
ggplot(subset(gtbl, country == "United States"),
       aes(x = year, y = lifeExp)) + geom_line() + geom_point()

```

Let just look at five countries

```{r}
jCountries <- c("United States", "Canada", "Rwanda", "Cambodia", "Mexico")
ggplot(subset(gtbl, country %in% jCountries),
       aes(x = year, y = lifeExp, color = country)) + geom_line() + geom_point()
```


So what's up with Mexico?

* Nafta? Higher GDP?

Not really...
```{r}
Countries <- c("United States", "Canada", "Rwanda", "Cambodia", "Mexico")
ggplot(subset(gtbl, country %in% jCountries),
       aes(x = year, y = lifeExp, color = country)) + geom_line() + geom_point(aes(size=gdpPercap))
```


## References

- Hadley Wickham's tutorials from useR 2014: Video [Part I](https://www.youtube.com/watch?v=8SGif63VW6E), [Part II](https://www.youtube.com/watch?v=Ue08LVuk790), [Slides and Tutorial](https://www.dropbox.com/sh/i8qnluwmuieicxc/AAAgt9tIKoIm7WZKIyK25lh6a).

- [Data Carpentry](http://datacarpentry.org/) (tidy data)[https://github.com/datacarpentry/datacarpentry/tree/master/lessons/tidy-data] lesson.

- The vignettes for [dplyr](http://cran.rstudio.com/web/packages/dplyr/) and [tidyr](http://cran.rstudio.com/web/packages/dplyr/)

* * *

- Derived from Jennifer Bryan, "hello ggplot2!", https://github.com/jennybc/ggplot2-tutorial/blob/master/ggplot2-tutorial-slides.pdf. License: CC-BY-NC
- Derived from Karthik Ram, "A quick introduction to ggplot2", [Speakerdeck](http://inundata.org/2013/04/10/a-quick-introduction-to-ggplot2/), [github](https://github.com/karthik/ggplot-lecture). License: [CC-BY](http://creativecommons.org/licenses/by/2.0/)


<!--  LocalWords:  'Lab Manipulation' html dplyr summarise frac gd df
 -->
<!--  LocalWords:  ggplot2 gapminder url gdf delim str tbl gtbl Stata
 -->
<!--  LocalWords:  lifeExp magrittr gdp gdpPercap usa gdpPercapRel gg
 -->
<!--  LocalWords:  NAFTA geoms aes colour ggplot log10 workflow 'p'
 -->
<!--  LocalWords:  lwd se lm Facetting gDat
 -->
