---
title: "Lab 6 - More Regression Interpretation"
author: "Jeffrey Arnold"
date: "Friday May 9, 2015"
---
This lab will use some libraries you've seen before, and one you may not have. We'll load them all now.

```{r, results = 'hide', echo = FALSE}
knitr::opts_chunk$set(tidy = TRUE,
                      tidy.opts = list(width.cutoff = 60, indent = 2))
```

```{r, message=FALSE, warning=FALSE}
library("MASS")
library("dplyr")
library("ggplot2")
library("broom")
```


```{r, results='hide'}
rossdata <- read.csv("http://pols503.github.io/pols_503_sp15/data/ross_2012.csv",
                     stringsAsFactors = FALSE)
```

## Regression adding coefficient

## Regression Post-estimation



## Writing Functions

This is an example of a really stupid R function that adds two to each variable.
```{r}
add2 <- function(x) {
  x + 2
}
```
Try it ...

This function creates a function that estimates 

Edit this function to create on that runs a regression on the rossoil dataset for a particular year. 
```{r}
function(yyyy) {
  filter()
}
```

Something like this would work.
You *need* to have the argument have a different name than the variable in data. 

Try 


## Sources

- <http://staff.washington.edu/csjohns/503/lab5.r>
- <http://staff.washington.edu/csjohns/503/lab6.r>
